%!TEX root =  main.tex
\section{System model and definitions}
\label{sec:sysmodel}

In this section, we present the system model, and define our
correctness criterion (i.e., linearizability). Additionally, because
\dynastar relies on two variations of a multicast primitive for
communication, we discuss them below. When the protocol requires that
commands be consistently ordered across partitions, \dynastar uses
atomic multicast. Otherwise, it uses a reliable multicast, which has
less overhead.


%\red{How do we use reliable multicast, as opposed to atomic?}
%\red{Do we need this primitive? Or is it just inherited from the artifact?}
%\lle{Partitions (and oracle) in \dynastar use reliable multicast to exchange
% objects and signals }
%% S-SMR relies on an atomic multicast primitive to consistently order
%% commands within and across partitions.  Commands that access objects
%% located in a single partition (i.e., single-partition commands) are
%% multicast to the concerned partition and executed like in classical
%% SMR.  Commands that access objects located in multiple partitions
%% (i.e., multi-partition commands) are multicast to all involved
%% partitions.  To prevent command interleaves that violate strong
%% consistency, partitions coordinate during the execution of
%% multi-partition commands.




\subsection{Processes and communication}

We consider a distributed system consisting of an unbounded set of
client processes $\ccm = \{c_1, c_2, ...\}$ and a bounded set of
server processes (replicas) $\ssm = \{s_1, ..., s_n\}$.  Set $\ssm$ is
divided into disjoint groups of servers $\ssm_0, ..., \ssm_k$.
Processes are either \emph{correct}, if they never fail, or
\emph{faulty}, otherwise.  In either case, processes do not experience
arbitrary behavior (i.e., no Byzantine failures).

Processes communicate by message passing, using either one-to-one or
one-to-many communication.  The system is asynchronous: there is no
bound on message delay or on relative process speed.  One-to-one
communication uses primitives $send(p,m)$ and $receive(m)$, where $m$
is a message and $p$ is the process $m$ is addressed to.  If sender
and receiver are correct, then every message sent is eventually
received.
%
One-to-many communication relies on reliable multicast and atomic
multicast,\footnote{Solving atomic multicast requires additional
  assumptions~\cite{CT96,FLP85}. In the following, we simply assume
  the existence of an atomic multicast oracle.}  defined in
sections~\ref{sec:rmcast} and \ref{sec:amcast}, respectively.


\subsection{Correctness criterion}
\label{sec:correctcrit}

Our consistency criterion is linearizability.  A system is
\emph{linearizable} if there is a way to reorder the client commands
in a sequence that (i)~respects the semantics of the commands, as
defined in their sequential specifications, and (ii)~respects the
real-time precedence of commands~\cite{Attiya04}.


\subsection{Reliable multicast}
\label{sec:rmcast}

To reliably multicast a message $m$ to a set of groups $\gamma$,
processes use primitive \rmcast$(\gamma, m)$.  Message $m$ is
delivered at the destinations with \rmdel$(m)$.  Reliable multicast
has the following properties:

\begin{itemize}

    \item[--] If a correct process \rmcast{}s $m$, then every correct
      process in $\gamma$ \rmdel{}s $m$ \emph{(validity)}.
    
    \item[--] If a correct process \rmdel{}s $m$, then every correct
      process in $\gamma$ \rmdel{}s $m$ \emph{(agreement)}.
    
    \item[--] For any message $m$, every process $p$ in $\gamma$
      \rmdel{}s $m$ at most once, and only if some process has
      \rmcast{} $m$ to $\gamma$ previously \emph{(integrity)}.
    
\end{itemize}

\subsection{Atomic multicast}
\label{sec:amcast}

To atomically multicast a message $m$ to a set of groups $\gamma$,
processes use primitive \amcast$(\gamma, m)$.  Message $m$ is
delivered at the destinations with \amdel$(m)$.  We define delivery
order $<$ as follows: $m < m'$ iff there exists a process that
delivers $m$ before $m'$.

Atomic multicast ensures the following properties:

\begin{itemize}
    
    \item[--] If a correct process \amcast{}s $m$, every correct
      process in a group in $\gamma$ \amdel{}s $m$ \emph{(validity)}.
    
    \item[--] If a process \amdel{}s $m$, then every correct process
      in a group in $\gamma$ \amdel{}s $m$ \emph{(uniform agreement)}.
    
    \item[--] For any message $m$, every process \amdel{}s $m$ at most
      once, and only if some process has \amcast{} $m$ previously
      \emph{(integrity)}.
    
    \item[--] The delivery order is acyclic \emph{(atomic order)}.

    \item[--] For any messages $m$ and $m'$ and any processes $p$ and
      $q$ such that $p \in g$, $q \in h$ and $\{ g, h \} \subseteq
      \gamma$, if $p$ delivers $m$ and $q$ delivers $m'$, then either
      $p$ delivers $m'$ before $m$ or $q$ delivers $m$ before $m'$
      \emph{(prefix order)}.
    
\end{itemize}

Atomic broadcast is a special case of atomic multicast in which there
is a single group of processes.

