%!TEX root =  main.tex

\clearpage
\section{Background}
\label{sec:background}

In this section, we first state the underlying assumptions about the system model (\S\ref{sec:system-model}) and then define the atomic multicast communication primitive (\S\ref{sec:amcast}).

\subsection{System model}
\label{sec:system-model}

We assume a hybrid distributed system model in which processes can use both message-passing and shared-memory \cite{Aguilera2019}.
The system is composed of a set of client and server processes, which communicate by exchanging messages or accessing portions of each other's memory.
Processes are \emph{correct}, if they do not fail, or \emph{faulty}, otherwise. 
In either case, processes do not experience arbitrary behavior (i.e., no Byzantine failures).
Our protocols ensure safety under both asynchronous and synchronous execution periods. 
To ensure liveness, we assume the system is \emph{partially synchronous} \cite{DLS88}, 
that is, it is initially asynchronous and eventually becomes synchronous. The time when the 
system becomes synchronous is called the \emph{Global Stabilization Time (GST)}, and it is unknown to the processes.
Before GST, there are no bounds on communication and processing delays; after GST, such bounds exist but are unknown. 

A process can share memory regions with other processes and define permissions for those shared memory regions. 
Process $q$ can read and write a register $v$ in $p$'s memory region $mr$ with operations $\textsf{read}(p,v)$ and $\textsf{write}(p,v,value)$, respectively.
A permission associated with memory region $mr$ defines disjoint sets of processes $R_{mr}, W_{mr}, RW_{mr}$ that can read, write, or read-write the registers in region $mr$. 
Process $q$ has permission to read (respectively, write and read-write) $v$ in $p$'s $mr$ if $q \in R_{mr}$ (resp. $W_{mr}, RW_{mr}$).
A process can initially assign permission for its shared memory regions and later change these permissions.

Processes can also communicate by exchanging messages over a set of directed links using primitives \textsf{send}$(p, m)$ and \textsf{receive}$(m)$, where $p$ is the addressee of message $m$.
We assume messages are unique.
Communication links are reliable in that every message sent by a process $p$ to another process $q$ is guaranteed to be eventually delivered by $q$ if both $p$ and $q$ are correct.
Moreover, a message is received at most once, and only if it was previously sent.


\subsection{Problem statement}
\label{sec:amcast}

Let $\Pi = \{p_1, p_2, ...\}$ be the set of server processes.
We define $\Gamma \in 2^{\Pi}$ as the set of process groups in the system. 
Groups are disjoint and each group contains $2f+1$ processes, where $f$ is the maximum number of faulty processes per group.  
The assumption of disjoint groups has little practical implication since it does not prevent collocating processes that are members of different groups on the same machine.
A set of $f + 1$ processes in group $g$ is a \emph{quorum} in $g$.

A process atomically multicasts a message $m$ to groups in $m.dst$ by invoking primitive multicast($m$), where $m.dst$ is a special field in $m$ with $m$'s destinations; a process delivers $m$ with primitive deliver($m$). 
We define the relation $<$ on the set of messages processes deliver as follows: $m < m'$ iff there exists a process that delivers $m$ before $m'$. 

Atomic multicast ensures the following properties: 
\begin{itemize}
\item \textit{validity}:~if a correct process $p$ multicasts a message $m$, then eventually all correct processes $q \in g$, where $g \in m.\mathit{dst}$, r-deliver $m$.
\item \textit{integrity}:~for any process $p$ and any message $m$, $p$ r-delivers $m$ at most once, and only if $p \in g$, $g \in m.\mathit{dst}$, and $m$ was previously r-multicast.
\item \textit{uniform agreement}:~if a process $p$ a-delivers a message $m$, then eventually all correct processes $q\in m.\mathit{dst}$ a-deliver $m$.
\item \textit{uniform prefix order}:~for any two messages $m$ and $m'$ and any two processes $p$ and $q$ such that $p \in g$, $q \in h$ and $\{ g, h \} \subseteq m.\mathit{dst} \cap m'.\mathit{dst}$, if $p$ a-delivers $m$ and $q$ a-delivers $m'$, then either $p$ a-delivers $m'$ before $m$ or $q$ a-delivers $m$ before $m'$.
\item \textit{uniform acyclic order}:~the relation $<$ is acyclic.
\end{itemize}
Atomic broadcast is a special case of atomic multicast in which every message is addressed to all groups.

We require atomic multicast protocols to be \emph{genuine}~\cite{GS01b}: 
an algorithm ${\cal A}$ solving reliable or atomic multicast is genuine
if and only if for any admissible run $R$ of ${\cal A}$ and for any process $p$ in $R$, if $p$ sends or receives a message, then some message $m$ is multicast, and either (a)~$p$ is the process that multicasts $m$ or (b)~$p \in g$ and $g \in m.dst$.






