%!TEX root =  main.tex
\section{Background}
\label{sec:background}

In this section, we present the system model (\S\ref{sec:system-model}), define the atomic multicast communication problem (\S\ref{sec:amcast}), and  overview RDMA, the key technology used by \libname (\S\ref{sec:rdma}).

\subsection{System model}
\label{sec:system-model}

We assume a hybrid distributed system model in which processes can use both message-passing and shared-memory \cite{Aguilera2019}.
The system is composed of a set of client and server processes.
Processes communicate by exchanging messages or accessing portions of each other's memory.
Processes are \emph{correct}, if they do not fail, or \emph{faulty}, otherwise. 
In either case, processes do not experience arbitrary behavior (i.e., no Byzantine failures).
Our protocols ensure safety under both asynchronous and synchronous execution periods. 
To ensure liveness, we assume the system is \emph{partially synchronous} \cite{DLS88}, 
that is, it is initially asynchronous and eventually becomes synchronous. The time when the 
system becomes synchronous is called the \emph{Global Stabilization Time (GST)}, and it is unknown to the processes.
Before GST, there are no bounds on communication and processing delays; after GST, such bounds exist but are unknown. 

A process can share memory regions with other processes and define permissions for those shared memory regions. 
Process $q$ can read and write a register $v$ in $p$'s memory region $mr$ with operations $\textsf{read}(p,v)$ and $\textsf{write}(p,v,value)$, respectively.
A permission associated with memory region $mr$ defines disjoint sets of processes, $R_{mr}, W_{mr}$, and $RW_{mr}$, that can read, write, and read-write the registers in region $mr$. 
Process $q$ has permission to read (respectively, write and read-write) $v$ in $p$'s $mr$ if $q \in R_{mr}$ (resp., $W_{mr}, RW_{mr}$).
A process can initially assign permissions for its shared memory regions and later change these permissions.

Processes can also communicate by exchanging messages over a set of directed links using primitives \textsf{send}$(p, m)$ and \textsf{receive}$(m)$, where $p$ is the addressee of message $m$.
We assume messages are unique.
Communication links are reliable in that every message sent by a process $p$ to another process $q$ is guaranteed to be eventually delivered by $q$ if both $p$ and $q$ are correct.
Moreover, a message is received at most once, and only if it was previously sent.


\subsection{Problem statement}
\label{sec:amcast}

Let $\Pi$ be the set of server processes in the system and $\Gamma \in 2^{\Pi}$ the set of process groups in the system, where $|\Gamma| = k$.
Groups are disjoint and each group contains $n = 2f+1$ processes, where $f$ is the maximum number of faulty processes per group.  
The assumption about disjoint groups has little practical implication since it does not prevent collocating processes that are members of different groups on the same machine.
Yet, it is important since atomic multicast requires stronger assumptions when groups intersect~\cite{GS01b}.
A set of $f + 1$ processes in group $g$ is a \emph{quorum} in $g$.

A process atomically multicasts a message $m$ to groups in $m.dst$ by invoking primitive multicast($m$), where $m.dst$ is a special field in $m$ with $m$'s destinations; a process delivers $m$ with primitive deliver($m$). 
We define the relation $<$ on the set of messages processes deliver as follows: $m < m'$ iff there exists a process that delivers $m$ before $m'$. 

Atomic multicast ensures the following properties: 
\begin{itemize}
\item \textit{Validity}:~if a correct process $p$ multicasts a message $m$, then eventually all correct processes $q \in g$, where $g \in m.\mathit{dst}$, deliver $m$.
\item \textit{Integrity}:~for any process $p$ and any message $m$, $p$ delivers $m$ at most once, and only if $p \in g$, $g \in m.\mathit{dst}$, and $m$ was previously multicast.
\item \textit{Uniform agreement}:~if a process $p$ delivers a message $m$, then eventually all correct processes $q\in m.\mathit{dst}$ deliver $m$.
\item \textit{Uniform prefix order}:~for any two messages $m$ and $m'$ and any two processes $p$ and $q$ such that $p \in g$, $q \in h$ and $\{ g, h \} \subseteq m.\mathit{dst} \cap m'.\mathit{dst}$, if $p$ delivers $m$ and $q$ delivers $m'$, then either $p$ delivers $m'$ before $m$ or $q$ delivers $m$ before $m'$.
\item \textit{Uniform acyclic order}:~the relation $<$ is acyclic.
\end{itemize}
Atomic broadcast is a special case of atomic multicast in which there is a single group (i.e., $\Gamma$ is a singleton).

We require atomic multicast protocols to be \emph{genuine}~\cite{GS01b}: 
an atomic multicast algorithm is genuine if in any run in which a message $m$ is 
multicast, then for every process $p$ that participates in ordering $m$, $p$ is the process that multicasts $m$ or $p \in g$ and $g \in m.dst$.

\subsection{Remote Direct Memory Access}
\label{sec:rdma}

Remote Direct Memory Access (RDMA) is a protocol that enables direct data access to the memory of a remote machine without involving the operating system and processor of the remote machine. 
RDMA implements the network stack in hardware, and provides both low latency and high bandwidth by bypassing the kernel and supporting zero-copy communication.
RDMA provides two-sided operations (e.g., send, receive), one-sided operations (e.g., read, write), and atomic operations (e.g., compare-and-swap, fetch-and-increment). The two-sided operations involve the CPU of the remote host and rely on user-space memory copies.
Thus, they introduce overhead when compared to one-sided RDMA verbs \cite{FaRM}.
Besides, previous studies have established that remote write operations provide performance superior to remote reads, and send and receive operations, and much better performance than atomic operations \cite{kalia2014using, kalia2016design, mitchell2013using}.
In RamCast's normal operation, processes communicate using remote write operations only.
We refrain from using remote read operations, and resort to send and receive operations when handling failures, since they lead to simpler logic.

RDMA provides three transport modes: Reliable Connection (RC), Unreliable Connection (UC) and Unreliable Datagram (UD). 
While RC and UC are connection-oriented and support only one-to-one data transmission, UD supports both one-to-one and one-to-many transmission without establishing connections.
RC ensures data transmission is reliable and correct in the network layer, while UC does not have such a guarantee.
In this work, we use RC to provide in-order reliable delivery.
To establish a connection between two remote hosts, the RDMA-enabled network card (RNIC) on each host creates a logical RDMA endpoint known as a Queue Pair (QP), including a send queue and a receive queue for storing data transfer requests. 
Operations are posted to QPs as Work Requests (WRs) to be consumed and executed by the RNIC. 
When an RDMA operation is completed, a completion event is pushed to a Completion Queue (CQ).
Operations can be made \emph{unsignaled} by setting a flag in the WR; these verbs do not generate a completion event, and the application detects completion using application-specific methods.
Each host makes local memory regions (MR) available for remote access by asking its OS to pin the memory pages that would be used by the RNIC.
Both QPs and MRs can have different access modes (i.e., read-only or read-write). 
The hosts specify the access mode when initializing the QP or registering the MR, but the access mode can be dynamically updated later. 
The host can register the same memory for different MRs. Each MR then has its own access mode.
In this way, different remote machines can have different access rights to the same memory region. 
% The same effect can be obtained by using different access flags for the QPs used to communicate with remote machines.