\begin{acknowledgements}

This thesis is the product of an interesting, albeit long and challenging,
journey. It would not be possible without the help of many people. I am
extremely grateful to all of them.

First and foremost, I would like to express my genuine gratitude to my advisor
Prof. Fernando Pedone, for the support of my Ph.D. study and research, and for
his inexhaustible patience, enthusiasm, and continuous encouragement. I have
learned many valuable lessons from his immense knowledge, from doing research,
presenting ideas, writing reports, and even windsurfing. Much of this
dissertation is a result of his endless support. I thank you.
  
I am grateful to the dissertation committee members, Miguel Correia, Robert
Soul\'e, Rachid Guerraoui, and Marc Langheinrich for their insightful comments
and the time they spent examining this thesis. In particular, I would like to
thank Paweł T. Wojciechowski for his detailed, constructive feedback and also
for the great time he hosted me in Poznań.
  
It has always been a privilege to be a part of the Distributed Systems Lab. I
would like to thank the current and former members of the group that I had the
opportunity to spend time with: Eduardo, Paulo, Edson, Daniele, Loan, Tu,
Mojtaba, Odorico, Pietro, Samuel, Theo, Tu, Nenad, Elia, Leandro and Fynn. I
would never forget those coffee times that we had together.
  
I am grateful to my big family for encouraging me and give unconditional support
throughout my whole life, especially my parents for their patience and love.
Lastly, I am and will always be thankful to my little family for standing with
me during last five year and be a part of some of the happiest moments in my
life.

\end{acknowledgements}
